### 2012-03-21 inst/doc/intro2fitdistrplus.Rnw
correction of the problem with optim using "L-BFGS-B" and new version of the vignette

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Copyright (c) 2009 Marie Laure Delignette-Muller, Christophe Dutang                        
%                                                                                                                                                       
%    This program is free software; you can redistribute it and/or modify                               
%    it under the terms of the GNU General Public License as published by                        
%    the Free Software Foundation; either version 2 of the License, or                                 
%    (at your option) any later version.                                                                                        
%                                                                                                                                                      
%    This program is distributed in the hope that it will be useful,                                          
%    but WITHOUT ANY WARRANTY; without even the implied warranty of                        
%    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                   
%    GNU General Public License for more details.                                                                    
%                                                                                                                                                         
%    You should have received a copy of the GNU General Public License                           
%    along with this program; if not, write to the                                                                            
%    Free Software Foundation, Inc.,                                                                                             
%    59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                            
%                                                                                                                                                         
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Fit parametric distributions on non-censored or censored data with fitdistrplus
%%%
%%%         Sweave vignette file
%%%

\documentclass[a4paper]{article}
% sweave commands for vignette
%\VignetteIndexEntry{Fit parametric distributions on non-censored or censored data}
%\VignettePackage{fitdistrplus}
%\VignetteKeyword{distribution}

% accents 8 bits dans le fichier
%\usepackage[applemac]{inputenc} %MAC encoding
%\usepackage[utf8]{inputenc} %UNIX encoding
\usepackage[ansinew]{inputenc} %WINDOWS encoding

\usepackage[a4paper,textwidth=18cm,textheight=27cm]{geometry}
\usepackage{url}

\title{Fitting parametric univariate distributions  to non censored or
censored data using the R \texttt{fitdistrplus} package}
\author{Marie Laure Delignette-Muller and Christophe Dutang}

\begin{document}
\setkeys{Gin}{width=0.6\textwidth}

\maketitle

<<echo=FALSE,results=hide>>=
options(digits=3)
@

% A enlever dans version JSS !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
\tableofcontents

%\pagebreak

\section{Introduction}
\label{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Overview}
\label{Overview}
%%%%%%%%%%%

Fitting distributions to data is a very common task in statistics. It consists in choosing
a probability distribution that gives a good representation of a statistical
variable. It requires judgment and expertise and generally needs an iterative process of
distribution choice, parameter estimation, and quality of fit evaluation. 
Function \texttt{fitdistr} in the R package \texttt{MASS} \cite{MASS} 
is a well known general-purpose maximum-likelihood
fitting routine for the parameter estimation step in R. Other steps of
the process may be developed using R \cite{Ricci05}. Our first objective by developping
package \texttt{fitdistrplus} \cite{fitdistrplus} 
was to provide R users a set of functions dedicated to help the overall
process of fitting a univariate parametric distribution to data.

Function \texttt{fitdistr} estimates distribution parameters by maximizing the log-likelihood
using function \texttt{optim}. In some cases, other estimation methods could be prefered, 
such as maximum goodness-of-fit estimation also commonly called minimum distance estimation, and
proposed in package \texttt{actuar} 
%\cite{actuar}
with three different goodness-of-fit distances. While developping
package \texttt{fitdistrplus}, our second objective  was to extend function \texttt{fitdistr} 
by providing various
estimation methods to fit distributions in addition to maximum likelihood. Functions
were developped to enable matching moment estimation, matching quantile estimation, and 
maximum goodness-of-fit estimation (or minimum distance estimation) using eight different distances.
Moreover, package \texttt{fitdistrplus} offers the possibility to specify a user-supplied function
for optimization, 
useful in cases where optimization techniques not included in function \texttt{optim} may be more adequate.

In applied statistics, it is not uncommon to have to fit distributions to censored data. 
Function \texttt{fitdistr} 
does not enable maximum likelihood estimation from this type data. Some packages 
deal with censored data, especially survival data \cite{survival},
but those packages generally focused on specific models, enabling the fit of only one distribution or a 
restricted family of distributions. Our third objective was thus to provide R users a function 
to estimate univariate distribution parameters from censored data, whatever the type of censoring.

% A changer en cas de mise a jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
This manuscript reviews the various features of version 0.3-4 of \texttt{fitdistrplus}. The package
is available from the Comprehensive R Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
The development version of the package is located at R-forge as one the packages of the 
project ``Risk Assessment with R''
(\url{http://r-forge.r-project.org/projects/riskassessment/})
The following command will load the package.
<<results=hide>>=
library(fitdistrplus)
@

\subsection{Running examples}
\label{Examples}
%%%%%%%%%%%
For illustrating the use of various functions of package \texttt{fitdistrplus}, we will use
four examples published in various biological areas, 
corresponding to data sets included in the package.

The two first data sets correspond to the observation of a continuous variable on a random sample
of a population of interest.

The ``ground beef'' data set contains values of
serving sizes in grams, collected in a French survey, for ground beef patties consumed by 
children under 5 years old. This data set was used in a quantitative risk assessment published
in a food microbiology journal (\cite{Delignette08}).
<<echo=TRUE>>=
data(groundbeef)
str(groundbeef)
@

The ``endosulfan'' data set contains
acute toxicity values for the organochlorine pesticide endosulfan
(geometric mean of LC50 ou EC50 values in $\mu g.L^{-1}$),
tested on Australian and non-Australian laboratory-species 
(arthropods, fish or nonarthropod invertebrates) (\cite{Hose04}).
<<echo=TRUE>>=
data(endosulfan)
str(endosulfan)
@

The ``Toxocara'' data set corresponds to the observation of a discrete variable, 
the number of \emph{Toxocara cati} parasites present in digestive tract,
on a random sample of feral 
cats living on Kerguelen island (\cite{Fromont01}).
<<echo=TRUE>>=
data(toxocara)
str(toxocara)
@

The ``smoked fish'' data set corresponds to the observation of a continuous censored variable, 
the \emph{Listeria monocytogenes} microbial concentration, on a random
sample of smoked fish distributed on the Belgian market in the period
2005 to 2007 (\cite{Busschaert10}).
Censored data are coded within 2 columns named left and right, describing
each observed value of \emph{Listeria monocytogenes} concentration 
(in $CFU.g^{-1}$) as an interval. 
The left column contains either \texttt{NA} for
left censored observations, the left bound of the interval for interval censored
observations, or the observed value for non-censored observations. The right
column contains either \texttt{NA} for right censored observations, the right bound of
the interval for interval censored observations, or the observed value for noncensored
observations.
<<echo=TRUE>>=
data(smokedfish)
str(smokedfish)
@

\section{Choice of candidate distributions}
\label{Choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Before fitting one or more distributions to a data set, it is generally
necessary to choose good candidates among a predefined family of distributions.
To help the user in this preliminary task, we developed functions to plot and
characterise empirical distributions.

\subsection{Graphical display of the observed distribution}
\label{Graphical}
%%%%%%%%%%

First of all, an empirical distribution may be plotted using classical R function
or using Function \texttt{plotdist} which provides plots
in density and in cdf as done in (Figure~\ref{plotdistcont}) for a continuous variable:

<<echo=TRUE, fig=FALSE>>=
plotdist(groundbeef$serving)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(groundbeef$serving)
@
  \caption{Density and cdf plots of an empirical distribution for a continuous variable 
  (serving size from the ``ground beef'' data set)}
  \label{plotdistcont}
\end{figure}

In some cases a discrete variable may be plotted as a continuous one, for example for a large data set
from a binomial distribution converging to a normal one, but
Function \texttt{plotdist} also proposes specific plots in density and in cdf
for discrete variables (Figure~\ref{plotdistdisc}):

<<echo=TRUE, fig=FALSE>>=
plotdist(toxocara$number,discrete = TRUE)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(toxocara$number,discrete = TRUE)
@
  \caption{Density and cdf plots of an empirical distribution for a discrete variable
  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{plotdistdisc}
\end{figure}

%\clearpage

\subsection{Empirical basis for selecting candidate distributions}
\label{Empirical}
%%%%%%%%%% 
Descriptives statistics may help the choice of good candidates to describe an empirical distribution
among a family of parametric distributions. Especially the skewness and kurtosis are useful for this
purpose. The concept of skewness relates to deviations from symmetry of the distribution.
The normal distribution has a skewness of zero. A positive (resp. negative) skewness indicates 
that the right (resp. left) tail of the distribution is more extended than the left (resp. right) one.
The concept of kurtosis relates to the tail weight. The normal distribution has a kurtosis of 3.
Distributions with a higher kurtosis are said to be leptokurtic, with heavier tails, such as the 
logistic distribution, while distributions with a smaller kurtosis are said platykurtic, with lighter 
tails, such as the uniform distribution.

Function \texttt{descdist} provides calculations of classical descriptive statistics
(minimum, maximum, median, mean, sample sd) and by default 
unbiased estimations of skewness and 
Pearsons's kurtosis values. Nevertheless estimations of skewness and kurtosis
are unbiased only for normal distributions and estimated values are thus only indicative.
A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is also provided for the 
empirical distribution (Figure~\ref{Cullenplotcont}). 
On this plot, values for common distributions are displayed as tools 
to help the choice of distributions to fit to data. For some distributions (normal, uniform,
logistic, exponential for example), there is only one possible value for the skewness and the kurtosis
 and the distribution is thus represented by a point on the plot. For other distributions, 
areas of possible values are represented, consisting in lines (as for gamma and lognormal distributions), 
or larger areas (as for beta distribution).

Skewness and kurtosis are known not to be robust. In order to take into account the uncertainty 
of the estimated values of kurtosis and skewness from data, the data set may be boostraped by 
fixing the argument \texttt{boot} to an integer above 10. Values of skewness and kurtosis 
corresponding to bootstrap samples are then computed and reported on the 
skewness-kurtosis plot. Below is a call to function \texttt{descdist} to describe the distribution 
of the serving size from the ``ground beef'' data set and to draw the corresponding skewness-kurtosis
plot (Figure~\ref{Cullenplotcont}). Looking at the results on this example
with a positive skewness and a kurtosis not far from 3,
the fit of three common right-skewed distributions could be considered, Weibull, gamma and 
lognormal distributions.

<<echo=TRUE, fig=FALSE>>=
descdist(groundbeef$serving,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(groundbeef$serving,boot=1000)
@
  \caption{Skewness-kurtosis plot for a continuous variable 
  (serving size from the ``ground beef'' data set)}
  \label{Cullenplotcont}
\end{figure}

For discrete variables, such as the number of \emph{Toxocara cati} parasites from 
the ``Toxacara'' data set,  skewness and kurtosis values or set of values of Poisson and negative binomial
distributions are represented in the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}), 
together with values for the
normal distribution, to which discrete distributions may converge. 
Looking at the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}) obtained
for the number of \emph{Toxocara cati} parasites from 
the ``Toxacara'' data set, one could try the fit of Poisson and negative-binomial distributions.

<<echo=TRUE, fig=FALSE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
  \caption{Skewness-kurtosis plot for a discrete variable
  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{Cullenplotdisc}
\end{figure}

\clearpage

\section{Fit of a distribution by maximum likelihood estimation}
\label{FIT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter estimation}
\label{Estim}
%%%%%%%%%%
Once selected, one or more parametric distributions may be fitted to the data set, one at a time, using
Function \texttt{fitdist}. By default, distribution parameters $\theta$ are estimated by maximizing the
likelihood defined as:
  \begin{equation}
    \label{likelihood}
    L(\theta)=\prod_{i=1}^n f(y_{i}|\theta)
  \end{equation}
with $y_{i}$ the $n$ observations of variable $y$ and $f$ the density function of the fitted 
parametric distribution.

% TO DO : change the formalism of this equation

The other proposed estimation methods are described in Section~\ref{Alternatives}.

Function \texttt{fitdist} returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted (see Figure~\ref{plotcontfit} in Section~\ref{GOF plots}). 
The parametric distribution must be a classically defined R distributions,
with at least d, p and q functions respectively for the density cdf and quantile functions
(for example \texttt{dnorm}, \texttt{pnorm} and \texttt{qnorm} for the normal distribution).
The name of the fitted distribution is specified in the first argument by its classical 
abbreviation used as the second part of d, p and q functions (for example ``norm''
for the normal distribution. Numerical results returned by Function \texttt{fitdist} are 
parameter estimates with estimated standard errors computed from the estimate
of the Hessian matrix at the maximum likelihood solution, correlation matrix
between parameter estimates, the loglikelihood, the Akaike and the Schwarz information criteria 
(so called AIC and BIC).
Below is a call to function \texttt{fitdist} to fit a Weibull distribution 
to the serving size in the ``ground beef'' data set.
<<echo = TRUE,fig = FALSE>>=
fw <- fitdist(groundbeef$serving,"weibull")
print(fw)
summary(fw)
@

The same procedure is required to fit a discrete distribution. As an example, 
using ``toxocara'' data set, Poisson and
negative distributions may be easily fitted and AIC values compared, in this case
giving the preference
to the negative binomial distribution, with a much smaller AIC value.
<<echo = TRUE,fig = FALSE>>=
(fp <- fitdist(toxocara$number,"pois"))
(fnb <- fitdist(toxocara$number,"nbinom"))
fp$aic
fnb$aic
@

For some distributions (see the help of \texttt{fitdist} for details), it is necessary to specify initial 
values for the distribution parameters in the argument \texttt{start} when using the maximum likelihood method.
\texttt{start} must be a named list
of parameters initial values. The names of
the parameters in \texttt{start} must correspond exactly to their definition in R or in a user-supplied R code.
Function \texttt{plotdist} (see Section~\ref{GOF plots}), 
which can plot any parametric distribution with specified parameter values
in argument \texttt{para} may help to find correct initial values for 
the distribution parameters in non trivial cases, by iterative calls if necessary (see~\cite{fitdistrplus}
for examples).

% Faudra trouver un autre exemple ou une autre distribution ou enlever cette partie

%As an example, here is the fit of a user-supplied distribution, the Gumbel distribution 
%(also named extreme value distribution) to an arbitrary data set, preceded by 
%a plot of this data set with the Gumbel distribution which parameters 
%were fixed at initial values manually obtained after iterative plots.

%<<echo=TRUE,fig=TRUE>>=
%x1 <- c(6.4, 13.3, 4.1, 1.3, 14.1, 10.6, 9.9, 9.6, 15.3, 22.1,
%13.4, 13.2, 8.4, 6.3, 8.9, 5.2, 10.9, 14.4)
%dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
%pgumbel<-function(q,a,b) exp(-exp((a-q)/b))
%qgumbel<-function(p,a,b) a-b*log(-log(p))
%plotdist(x1,"gumbel",para=list(a=10,b=5))
%fitdist(x1,"gumbel",start=list(a=10,b=5))
%@
%\begin{figure}[htb]
%  \centering
%<<echo=FALSE, results=hide, fig=TRUE>>=
%plotdist(x1,"gumbel",para=list(a=10,b=5))
%@
%  \caption{Preliminar plot of a distribution against data to find initial values for parameters}
%  \label{plotdist}
%\end{figure}

\subsection{Goodness-of-fit plots}
\label{GOF plots}
%%%%%%%%%%
The plot of an object of class \texttt{fitdist} provides two types of results depending of the nature 
of the distribution, continuous or discrete. For continuous distributions, 
four goodness-of-fit plots are provided : a draw of pdf curve and histogram together, an cdf plot
of both empirical and theoretical distributions, a Q-Q plot
(plot of the quantiles of the theoretical fitted distribution (x-axis) against the empirical quantiles of the data)
and a P-P plot (i.e. for each value of the data set, plot of the cumulative density function of the fitted distribution
(x-axis) against the empirical cumulative density function (y-axis)) are also given (\cite{Cullen99}).
The Q-Q plot emphasizes the lack-of-fit at the distribution tails while the P-P plot emphasizes
the lack-of-fit at the distribution center.
As an example, let us look at the plot of the previous fit of a Weibull distribution to the``groundbeef'' data set
(Figure~\ref{plotcontfit}). The fit is not perfect,
especially in the center of the distribution, but seems correct while looking at the tails.

<<echo=TRUE, fig=FALSE>>=
plot(fw)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fw)
@
  \caption{Plot of the fit of a continuous distribution (a Weibull
  distribution fitted to serving sizes from the ``ground beef'' data set)}
  \label{plotcontfit}
\end{figure}

For continuous distributions, Function \texttt{cdfcomp} enables the visual comparison
of empirical and theoretical cumulative distributions for various distributions fitted
on a same data set. Function \texttt{cdfcomp} must be called with a first argument 
corresponding to a list of objects of class \texttt{fitdist}, and optionaly
further arguments to customize the plot, as in the following example comparing the fit of
Weibull, lognormal and gamma distributions to ``groundbeef'' data set (Figure~\ref{compcdf}).
<<echo=TRUE, fig=FALSE>>=
    fg <- fitdist(groundbeef$serving,"gamma")
    fln <- fitdist(groundbeef$serving,"lnorm")
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
  \caption{Comparison of CDF plots of various distributions fitted on continuous data
  (Weibull, gamma and lognormal
  distributions fitted to serving sizes from the ``ground beef'' data set)}
  \label{compcdf}
\end{figure}

In such a plot, data may be represented in a log scale when required, by just fixing the argument
\texttt{xlogscale} to \texttt{TRUE} in the call to \texttt{cdfcomp}.
% Ajouter ex endosulfan avec la bibliothËque actuar !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% mais comment faire dans la vignette : indiquer que fitdistrplus nÈcessite actuar????????????????????????????

For discrete distributions, the plot of an object of class \texttt{fitdist} provides
two goodness-of-fit plots comparing empirical and theoretical distributions
in pdf and in cdf.
As an exemple, let us look at the plot of the previous fit of a negative binomial distribution 
to ``toxocara'' data set.

<<echo=TRUE, fig=FALSE>>=
plot(fnb)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fnb)
@
  \caption{Plot of the fit of a discrete distribution (a negative binomial distribution
  fitted to numbers of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{plotdiscfit}
\end{figure}

\subsection{Measures of goodness-of-fit}
\label{GOF measures}
%%%%%%%%%%
When fitting continuous distributions, 
Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}
as defined by Stephens (\cite{Stephens86}).
<<echo=TRUE,fig=FALSE>>=
gofstat(fw)
@

% ADD the definition of these statistics !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%EQUATIONS TO BE ADDED

As giving  more weight to distribution tails, 
Anderson-Darling statistics is of special interest where it is important to place equal 
emphasis on fitting a distribution at the tails as well as the main body, 
as it is often the case in risk assessment \cite{Cullen99,Vose10}. 
% ADD AN EXAMPLE with comparison !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
Nevertheless, this statistics should 
be used cautiously when comparing fits of various distributions, keeping in mind that the weighting of each
cdf quadratic difference is dependent of the theoretical distribution.

When fitting discrete distributions, the Chi-squared statistic is computed by 
Function \texttt{gofstat} using cells defined 
by the argument \texttt{chisqbreaks} or cells automatically defined from the data in order 
to reach roughly the same number of observations per cell, roughly equal to the argument
\texttt{meancount}, or sligthly more if there are some ties. 
The choice to define cells from the empirical distribution (data) and not from the
theoretical distribution was done to enable the comparison of Chi-squared values obtained
with different distributions fitted on a same dataset.
If arguments \texttt{chisqbreaks} and \texttt{meancount} 
are both omitted, \texttt{meancount} is fixed in order to obtain roughly $(4n)^{2/5}$ cells, 
with $n$ the length of the dataset~\cite{Vose10}. Using this default option with 
the fit of a negative binomial distribution to ``toxocara'' data set gives following results :
<<echo=TRUE,fig=FALSE>>=
gofstat(fnb)
@

Among its returned values, Function \texttt{gofstat} provides a table with observed and theoretical 
counts used for the Chi-squared calculations:
<<echo=TRUE,fig=FALSE>>=
gofstat(fnb)$chisqtable
@

Even if specifically recommended for discrete distributions, the Chi-squared statistic may also be used for
continuous distributions (see~\cite{fitdistrplus} for examples).

\subsection{Goodness-of-fit tests}
\label{GOF tests}
%%%%%%%%%%

For continuous distributions, an approximate Kolmogorov-Smirnov test is 
performed by assuming the distribution parameters known. The critical value defined by Stephens~\cite{Stephens86} 
for a completely specified distribution is used to reject or not the 
distribution at the significance level 0.05. Because of this approximation, the result of the test
(decision of rejection of the distribution or not) is returned only for datasets with more 
than 30 observations. Note that this approximate test may be too conservative.

For datasets with more than 5 observations and for continuous distributions for 
which the test is described by Stephens~\cite{Stephens86} (normal, lognormal,
exponential, Cauchy, gamma, logistic and Weibull),
the Cramer-von Mises and Anderson-darling tests are performed as described by Stephens~\cite{Stephens86}. 
Those tests take into 
account the fact that the parameters are not known but estimated from the data. The result is the 
decision to reject or not the distribution at the significance level 0.05. Both tests are available
only for maximum likelihood estimations.

When the Chi-squared statistic is computed (for discrete or optionnaly continuous distributions), 
and if the degree of freedom (nb of cells - nb of parameters - 1)  of the corresponding distribution 
is strictly positive, the p-value of the Chi-squared test is returned.

Goodness-of-fit tests may be used carefully. As for any null-hypothesis significance
test, the non reject of the null hypothesis dose not imply its acceptation. However, this
misinterpretation of p-values is very common and comes from the wrong assumption that
absence of evidence is evidence of absence ~\cite{Altman96}.
On the contrary, in some cases, especially on very big datasets, 
even if the null hypothesis is rejected, a fitted distribution may be 
chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
plots do not show strong differences between empirical and theoretical distributions.

% ADD an example to argument that !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

Now let us look at the Chi-squared test results for the fit of a negative binomial distribution 
to ``toxocara'' data set :
<<echo=TRUE,fig=FALSE>>=
gofstat(fnb,print.test = TRUE)
@

A warning message appears as one of the theoretical counts is under 5 using the default breaks 
(see Section~\ref{GOF measures}). In order to solve this problem, one may specify breaks 
more adapted for the realization of the test.
<<echo=TRUE,fig=FALSE>>=
gofstat(fnb,chisqbreaks=c(0,1,4,8,20),print.test=TRUE)$chisqtable
@

From goodness-of-fit graphs, Chi-squared statistics, AIC and BIC values,
it seems better to choose the fit of a negative binomial distribution
for this dataset even it has one more parameter than the Poisson one. This was not obvious while
looking at the skewness-kurtosis graph. This graph must be used cautiously especially for continuous 
distributions far from the normal distribution or for discrete distributions. It is only indicative.

%\clearpage

\section{The special case of censored data}
\label{censored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Censored data may contain left censored, right censored and interval censored values, 
with several lower and upper bounds. Data must be coded into a dataframe with two columns,
 respectively named \texttt{left} 
 and \texttt{right}, describing each observed value as an interval.
 The \texttt{left} column contains either \texttt{NA} for left censored observations,
 the left bound of the interval for interval censored observations,
 or the observed value for non-censored observations.
 The \texttt{right} column contains either \texttt{NA} for right censored observations,
 the right bound of the interval for interval censored observations,
 or the observed value for non-censored observations.

\subsection{Graphical display of the observed distribution}
\label{censored graph}
%%%%%%%%%%

Using censored data such as those coded in the ``smokedfish'' data set,
the empirical distribution may be plotted using the function \texttt{plotdistcens}.
By default this function uses the EM approach of Turnbull \cite{Turnbull74} to compute the overall
empirical cdf curve with confidence intervals, by calls to functions \texttt{survfit}
and \texttt{plot.survfit} from the \texttt{survival} package.
Let us see such a plot for ``smokedfish'' data set after classical transformation
of microbial counts in decimal logarithm (Figure~\ref{plotdistcens}).

% ------------------ previous method -----------------------------
%Data are reported directly as segments for interval, left and right censored data, 
%and as points for non-censored data. Before plotting, observations are ordered and a rank r
%is associated to each of them. Left censored observations are ordered
%first, by their right bounds. Interval censored and non censored observations
%are then ordered by their mid-points and, at last, right censored observations are
%ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
%left censored (resp. right censored) observations are considered as interval censored
%observations and ordered by mid-points with non-censored and interval censored data.
%It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
%extreme value, even if not exactly known, to obtain a reasonable global ranking of 
%observations. After ranking, each of the n observations is plotted as a point (one x-value) 
%or a segment (an interval of possible x-values),
%with an y-value equal to r/n, r being the rank of each observation in the global ordering
%previously described.

<<echo=TRUE, fig=FALSE>>=
log10C <- data.frame(left=log10(smokedfish$left),right=log10(smokedfish$right))
plotdistcens(log10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plotdistcens(log10C)
@
  \caption{CDF plot of censored data (microbial counts from 
   the ``smoked fish'' data set)}
  \label{plotdistcens}
\end{figure}

\subsection{Maximum likelihood estimation}
\label{censored MLE}
%%%%%%%%%%

As for non censored data, one or more parametric distributions may then 
be fitted to the censored data set, one at a time, but using in this case 
Function \texttt{fitdistcens}. This function estimates the distribution parameters
by maximizing the likelihood given by following equation for censored data.
As \texttt{fitdist}, it returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted.
% ADD the equation !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

For ``smokedfish'' data set, a normal distribution may be fitted to log
transformed data as commonly done for microbial count data. 
<<echo=TRUE,fig=FALSE>>=
flog10C <- fitdistcens(log10C,"norm")
print(flog10C)
summary(flog10C)
@

As with \texttt{fitdist}, for some distributions (see \cite{fitdistrplus} for details), 
it is necessary to specify initial 
values for the distribution parameters in the argument \texttt{start}.
The function \texttt{plotdistcens} may help to find correct initial values for 
the distribution parameters in non trivial cases, by an manual iterative use if necessary.

\subsection{Goodness-of-fit plot}
\label{censored GOF graph}
%%%%%%%%%%
Only one goodness-of-fit plot is provided for censored data (Figure~\ref{plotfitdistcens}), corresponding to the theoretical 
cumulative distribution function added to the plot of censored data presented
in Section~\ref{censored graph}.

<<echo=TRUE, fig=FALSE>>=
plot(flog10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(flog10C)
@
  \caption{Goodness-of-fit CDF plot for a fit of a continuous distribution
  on censored data (a lognormal distribution fitted to microbial counts from 
   the ``smoked fish'' data set)}
  \label{plotfitdistcens}
\end{figure}

Computations of goodness of fit statistics have not yet been 
developed for fits using censored data, so the quality of fit 
may only be estimated from the loglikelihood and the goodness-of-fit CDF plot.

\section{Alternative methods for parameter estimation}
\label{Alternatives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Maximum goodness-of-fit estimation}
\label{MGE}
%%%%%%%%%%
Maximum likelihood is only the default estimation method proposed by Function
\texttt{fitdist}, but other methods may be used to estimate parameters for non-censored data.
One of the alternative for continuous distributions
is the maximum goodness-of-fit estimation method also called
minimum distance estimation method. In this package this method is proposed with
eight different distances the three classical distances defined in \cite{Stephens86}, 
(Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling 
which gives more weight to the tails of the distribution),
or one of the variants of this last distance proposed by \cite{Luceno06}. 
The right-tail AD gives more weight only to the right tail, the left-tail AD 
gives more weight only to the left tail. Either of the tails, or both of them, 
can receive even larger
weights by using second order Anderson-Darling Statistics.

%Computational forms of these distances are given in Table~\ref{MGEdist}.
% ADD TABLE with formulas !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

To fit a distribution by maximum goodness-of-fit estimation, one
needs to fix the argument \texttt{method}
to ``mge'' in the call to \texttt{fitdist} and to specify the argument \texttt{gof}
coding for the chosen goodness-of-fit distance. 
% Character string coding for the eight 
% proposed goodness-of-fit distances are given in Table~\ref{MGEdist}.
This function is intended to be used only with continuous variables and distributions.
It may be useful to fit distributions for which maximum likelihood does not provide
good estimations, such as the uniform distribution (\cite{Luceno06}).

<<>>=
u <- runif(50)
fitdist(u,"unif",method="mge",gof="KS")
@

Maximum goddness-of-fit estimation may also be useful to give
more weight to data at one tail of the distribution. In ecotoxicology, species sensitivity distributions such 
as those presented in \cite{Hose04} are often
fitted by a lognormal distribution (or another parametric distribution)
so as to estimate a low percentile, often 5$\%$ percentile,
named the hazardous concentration 5$\%$ (HC5). This value is then interpreted as a value of the
contaminant concentration protecting 95$\%$ of the species. In this context, one may consider
to fit the parametric distribution by giving more weight to the left tail of the empirical distribution
such as in the following example using left tail Anderson-Darling distances of first or second order
(Figure~\ref{plotfitMGE}).

<<echo=TRUE, fig=FALSE>>=
data(endosulfan)
ATV <-subset(endosulfan,group == "NonArthroInvert")$ATV
flnMGEKS <- fitdist(ATV,"lnorm",method="mge",gof="KS")
flnMGEAD <- fitdist(ATV,"lnorm",method="mge",gof="AD")
flnMGEADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
flnMGEAD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
xlogscale = TRUE,main="",
legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
xlegend = 500, ylegend = 0.15)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
xlogscale = TRUE,main="",
legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
xlegend = 500, ylegend = 0.15)
@
  \caption{Comparison of one distribution fitted by maximum goodness-of-fit using
  various goodness-of-fit distances (a lognormal distribution fitted to acute toxicity values
  from the ``endosulfan'' data set)}
  \label{plotfitMGE}
\end{figure}

\subsection{Moment matching estimation}
\label{MME}
%%%%%%%%%%
Moment matching estimation may also be performed fixing the argument \texttt{method}
to ``mme'' in the call to fitdist.
<<>>=
fitdist(u,"unif",method="mme")
@

The estimate is computed by a closed formula for following distributions:
normal, lognormal, exponential, Poisson, gamma, logistic, negative binomial,
geometric, beta and uniform distributions.
For distributions characterized by one parameter (geometric, Poisson and exponential),
this parameter is simply estimated by matching theoretical and observed means, and for distributions 
characterized by two parameters, these parameters are estimated by matching theoretical and observed 
means and variances (\cite{Vose10}).

For other distributions, Function \texttt{fitdist} carries out 
the matching numerically, by minimization of the
sum of squared differences between observed and theoretical moments (see \cite{fitdistrplus}
for technical details).

\subsection{Quantile matching estimation}
\label{QME}
%%%%%%%%%%
Quantile matching may also be performed fixing the argument \texttt{method}
to ``qme'' in the call to fitdist and adding an argument \texttt{probs}
defining the probabilities for which the quantile matching is performed.
The length of this vector must be equal
to the number of parameters to estimate. The quantile matching is carried out numerically, 
by minimizing the
sum of squared differences between observed and theoretical quantiles.
Here is an example of fit of a uniform distribution by matching first
and third quartiles.
<<>>=
fitdist(u,"unif",method="qme",probs=c(0.25,0.75))
@

\subsection{Customization of the optimization algorithm}
\label{Customization}
%%%%%%%%%%
Each time a numerical minimization (or maximization) is carried out using \texttt{fitdist}, 
Function \texttt{optim} of the package \texttt{stats} is used by default with the 
``Nelder-Mead'' method for distributions characterized by more than one parameter and
the ``BFGS'' method for distributions characterized by only one parameter

Sometimes the default algorithm fails to converge. It may then
be interesting to change some options of the function \texttt{optim} or to use another optimization
function than \texttt{optim} to maximize the likelihood.

The argument \texttt{optim.method} may be used in the call to \texttt{fitdist} or
\texttt{fitdistcens}. It will internally be passed to \texttt{mledist} and to \texttt{optim}.
This argument may be fixed to ``Nelder-Mead'' (the robust Nelder and Mead method), 
``BFGS'' (the BFGS quasi-Newton method), ``CG'' (a conjugate gradients method), 
``SANN'' (a variant of simulated annealing) or ``L-BFGS-B'' (a modification of the BFGS 
quasi-Newton method which enables box constraints optimization). For the use of the last method the 
arguments \texttt{lower} and/or \texttt{upper} also have to be passed. More details on these optimization
functions may be found in the help page of \texttt{optim} from the package \texttt{stats}.

Here are examples of fits of a gamma distribution to ``ground beef'' data set with various options
of \texttt{optim}.

<<>>=
fitdist(groundbeef$serving,"gamma",optim.method="Nelder-Mead")
fitdist(groundbeef$serving,"gamma",optim.method="BFGS") 
fitdist(groundbeef$serving,"gamma",optim.method="SANN")
@

You may also want to use another function than \texttt{optim} to maximize the likelihood.
This optimization function has to be specified by the argument \texttt{custom.optim} 
in the call to \texttt{fitdist} or
\texttt{fitdistcens}. But before that, it is necessary to customize this optimization function :
\texttt{custom.optim} function must have (at least) the following arguments,
\texttt{fn} for the function to be optimized, \texttt{par} for the initialized parameters. 
It is assumed that \texttt{custom.optim} should carry out a MINIMIZATION. Finally, it should return
at least the following components: \texttt{par} for the estimate, \texttt{convergence} for the convergence
code, \texttt{value} for \texttt{fn(par)} and \texttt{hessian}.

Below is an example of code written to customize \texttt{genoud} function from \texttt{rgenoud} package.
\begin{verbatim}
mygenoud <- function(fn, par, ...) 
{
   require(rgenoud)
   res <- genoud(fn, starting.values=par, ...)        
   standardres <- c(res, convergence=0)
   return(standardres)
}
\end{verbatim}

The customized optimization function may then be passed as the argument \texttt{custom.optim} 
in the call to \texttt{fitdist}
or \texttt{fitdistcens}. The following code may for example be used to fit a gamma distribution to the
``ground beef'' data set. Note that in this example various arguments are also passed from \texttt{fitdist} 
to \texttt{genoud} :
\texttt{nvars}, \texttt{Domains}, \texttt{boundary.enforcement}, 
\texttt{print.level} and \texttt{hessian}.

\begin{verbatim}
fitdist(groundbeef$serving, "gamma", custom.optim=mygenoud, nvars=2,    
   Domains=cbind(c(0,0), c(10, 10)), boundary.enforcement=1, 
   print.level=1, hessian=TRUE)
\end{verbatim}

\section{Uncertainty in parameter estimates}
\label{Uncertainty}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bootstrap procedures}
\label{Bootstrap}
%%%%%%%%%%

The uncertainty in the parameters of the fitted distribution may be simulated by parametric or nonparametric
boostrap using the function \texttt{boodist}. This function returns the boostrapped values of parameters in a 
S3 class object which
may be plotted to visualize the bootstrap region. 
The medians and the 95 percent confidence intervals of parameters (2.5 and 97.5 percentiles) are printed 
in the summary.
If inferior to the whole number of iterations, the number of iterations for which the function converges 
is also printed in the summary.

The plot of an object of class \texttt{bootdist} consists in a scatterplot or a matrix of scatterplots
of the bootstrapped values of parameters providing
a representation of the joint uncertainty distribution of the fitted parameters (Figure~\ref{bootstrap}).

Below is an example of the use of this function with the previous of the Weibull distribution to 
``groundbeef`` data set. 
<<echo=TRUE, fig=FALSE>>=
bw<-bootdist(fw,niter=1001)
plot(bw)
summary(bw)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(bw)
@
  \caption{Bootstrappped values of parameters for a fit of a distribution
characterized by two parameters (example on the fit of a Weibull distribution to
serving sizes from the ``ground beef'' data set)}
  \label{bootstrap}
\end{figure}

Function \texttt{boodistcens} provides the same type of results for fit on continuous distributions to
censored data.
<<echo=TRUE, fig=FALSE>>=
blog10C<-bootdistcens(flog10C,niter=1001)
summary(blog10C)
@

% ADD an example with the Burr distribution characterized by three parameters

%\subsection{Bootstrap confidence intervals and regions}
%\label{confidence}
%%%%%%%%%%

%To articulate with previous part

\subsection{Use of bootstrap samples}
%\label{MC2D}
%%%%%%%%%%
%TO BE DONE
Bootstrap samples of parameter estimates may be used to calculate confidence intervals on each parameter
of the fitted distribution, but it is also interesting to look at the marginal distribution of the bootstrap
values in a scatterplot (or a matrix of scatterplots if the number of parameters exceeds two), and especially
to look at the potential structural correlation between parameters.

Moreover, the use of the whole bootstrap sample is of great interest in the risk assessment field. Its use enables
the characterization of uncertainty in distribution parameters. It can be directly used within a second order Monte Carlo
simulation framework, especially within the package \texttt{mc2d} (\cite{mc2d}). One could refer to Pouillot \emph{et al.}
(\cite{Pouillot10}) for an introduction to the use of \texttt{mc2d} 
and \texttt{fitdistrplus} packages in the context of quantitative 
risk assessment.

\clearpage

%\section{Conclusion}
%\label{ccl}
%%%%%%%%%%
%TO BE DONE

\bibliographystyle{plain}
\bibliography{intro2fitdistrplus}

\end{document}