### 2020-05-15 share/prez/2018recontresR/Delignette_slides4R2108-v7.Rmd

---
title: "New goodness-of-fit plots for censored data in the package **fitdistrplus**"
author: |
 | M.L. Delignette-Muller (LBBE, Lyon)
 | C. Dutang (CEREMADE, Paris) and 
 | A. Siberchicot (LBBE, Lyon)
date: "July 6th 2018"
output:
  beamer_presentation:
 # theme: "Antibes"
 #   toc: true
    fig_width: 8
    fig_height: 6
    fig_caption: false
---
```{r, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
options(digits = 3)
require(fitdistrplus)
set.seed(123)
addbounds <- function(d)
{
  xbounds <- c(d$left, d$right)
  xboundsnotNA <- xbounds[!is.na(xbounds)]
  abline(v = xboundsnotNA, col = "grey")
}
addLR <- function(d)
{
  Lbounds <- d$left[!is.na(d$left)]
  Rbounds <- d$right[!is.na(d$right)]
  range <- range(c(Lbounds,Rbounds)) 
  eps <- (range[2] - range[1]) * 0.01
  text(x = Lbounds-eps, y = 0.05, labels = "L", col = "red")
  text(x = Rbounds+eps, y = 0.05, labels = "R", col = "red")
}

addeq <- function(d)
{
  left <- d$left
  left[is.na(left)] <- -100
  right <- d$right
  right[is.na(right)] <- 100
  rect(left, -2, right, 2, density = 10)
}

addTurnbull <- function(d, conf.int = FALSE)
{
  (survdata <- Surv(time = d$left, time2 = d$right, type="interval2"))
  survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
  points(survfitted$time, 1 - survfitted$surv, col = "orange", type = "s", 
         lwd = 5, lty = 2)
  if (conf.int == TRUE)
  {
      points(survfitted$time, 1 - survfitted$lower, col = "orange", type = "s", 
             lwd = 2, lty = 3)
      points(survfitted$time, 1 - survfitted$upper, col = "orange", type = "s", 
             lwd = 2, lty = 3)
  }
}
```

## General presentation of the package fitdistrplus
A package to help the **fit of parametric distributions** to 
univariate discrete our continuous non censored or censored data.

* stable version 1.0-9 on **CRAN** (first release in 2009).
* version 1.0-10 in development on **Rforge** (soon on CRAN).
* Delignette-Muller, M. L., & Dutang, C. (2015). fitdistrplus: An R package for fitting distributions. **Journal of Statistical Software**, 64(4), 1-34. (311 citations in scholar google)
* A FAQ vignette continuously updated in each new version.

```{r, echo = FALSE, fig.height=3}
par(mfrow = c(1,3))
set.seed(124)
r <- rgamma(100, shape = 8, rate = 1)
f <- fitdist(r, "gamma")
b <- bootdist(f)
f.mme <- fitdist(r, "gamma", method = "mme")
f.mge <- fitdist(r, "gamma", method = "mge", gof="AD2R")
f.qme <- fitdist(r, "gamma", method = "qme", probs = c(0.25, 0.75))
denscomp(list(f, f.mme, f.qme, f.mge), legendtext = c("MLE", "MME", "QME", "MGE"),
         main = "Different fitting methods", xlim = c(2, 16), fitlty = 1)
CIcdfplot(b, CI.output = "quantile", CI.fill = "pink", main = "Uncertainty using bootstrap")

rd <- rpois(50, lambda = 5)
fp <- fitdist(rd, "pois")
denscomp(fp, demp = TRUE, main = "Fit of discrete distributions")
```

# Goodness-of-fit plots for non censored data

## An example with non censored data
```{r}
r <- rweibull(100, shape = 3, scale = 1)
fw <- fitdist(r, "weibull")
fg <- fitdist(r, "gamma")
fl <- fitdist(r, "lnorm")
gofstat(list(fw, fg, fl), 
        fitnames = c("Weibull", "gamma", "lnorm"))
```

## A goodness-of-fit plot in density plot
```{r}
denscomp(list(fw, fg, fl), demp = TRUE, fitlty = 1)
```

## A goodness-of-fit plot in CDF
```{r}
cdfcomp(list(fw, fg, fl), fitlty = 1)
```

## A Q-Q plot which emphasizes differences at tails
```{r}
qqcomp(list(fw, fg, fl))
```

## a P-P plot which emphasizes differences in the center
```{r}
ppcomp(list(fw, fg, fl))
```

# Representation of the ECDF for censored data

## How to represent an ECDF from censored data ?

**A first toy example with left, right and interval censored data**
```{r, echo = FALSE}
d0 <- data.frame(left = c(NA, 2, 4, 7, 10), right = c(1, 2.5, 6, 8, NA))
d <- d0
```
```{r}
d
```
```{r, echo = FALSE, fig.height=2}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "") 
```

## Non Parametric Maximum Likelihood Estimation (NPMLE) of the ECDF: the Turnbull plot (package survival) used in former versions of fitdistrplus.
```{r, echo = FALSE,fig.height=5}
par(mar = c(3, 4, 0.1, 0.1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
plotdistcens(d, NPMLE.method = "Turnbull", lwd = 5, col = "orange", main = "")
addbounds(d)
```

## A new algorithm and plot from the package npsurv (Wang)
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

Equivalence classes on the first toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
addeq(d)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

Equivalence classes on a second toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
d0b <- data.frame(left = c(NA, 2, 4, 6, 10), right = c(1, 3, 7, 8, NA))
d <- d0b
deq <- data.frame(left = c(NA, 2, 6, 10), right = c(1, 3, 7, NA))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
addeq(deq)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

2. Assign a **probability mass** to each equivalence class (may be 0).

The Wang plot on the second toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addbounds(d)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

2. Assign a **probability mass** to each equivalence class (may be 0).

Various algorithms implemented in the packages **Icens**, **interval** and **npsurv** (more or less performant and not all handling left censored data).

## A third toy example
```{r, echo = FALSE}
d3 <- data.frame(left = c(-1.4, -1.4, 2, -1.4, 0),
                 right = c(1, 2, NA, 0, 2))
d <- d3
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The third toy example with the add of a non censored obs.
```{r, echo = FALSE}
d3b <- data.frame(left = c(1.18, -1.4, -1.4, 2, -1.4, 0),
                 right = c(1.18, 1, 2, NA, 0, 2))
d <- d3b
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## A realistic example: data salinity
```{r, echo = FALSE}
data(salinity)
d <-  log10(salinity)
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 0.2, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addbounds(d)
```

# New CDF, Q-Q and P-P plots implemented for censored data

## Use of cdfcompcens() to assess the fit of 3 distributions on data smokedfish
```{r, echo = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
dgumbel <- function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
pgumbel <- function(q,a,b) exp(-exp((a-q)/b))
qgumbel <- function(p,a,b) a-b*log(-log(p))
fitsfg<-fitdistcens(d,"gumbel",start=list(a=-3,b=3))
```
```{r, echo = FALSE, fig.height = 3.5, fig.width = 5}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2), nrow =3, byrow = TRUE))
par(mar = c(3, 4, 2, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 0.2, col = "blue", main = "")
# addbounds(d)
cdfcompcens(list(fitsfn,fitsfl,fitsfg), fitlty = 1)
# addbounds(d)
```

## Use of qqcompcens() for one distribution
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
qqcompcens(fitsfn)
```

## Use of ppcompcens() for one distribution
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
ppcompcens(fitsfn)
```

## Q-Q plots and P-P plot for the 3 distributions
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
qqcompcens(list(fitsfn,fitsfl,fitsfg))
ppcompcens(list(fitsfn,fitsfl,fitsfg))
```

## An alternative presentation of the Q-Q plots for the 3 dist. 
```{r, echo = FALSE, fig.height = 5.5, fig.width = 6}
par(mfrow = c(2,2))
par(mar = c(4, 4, 2, 1))
qqcompcens(fitsfn, fitcol = "red", fillrect = "red")
qqcompcens(fitsfl, fitcol = "green", fillrect = "green")
qqcompcens(fitsfg, fitcol = "blueviolet", fillrect = "blueviolet")
```

Will be soon implemented in the plotstyle ggplot.

## Another example with data salinity
```{r, echo = FALSE, fig.height = 5.5, fig.width = 6}
data(salinity)
d <-  log10(salinity)
f <- fitdistcens(d,"norm")
par(mfrow = c(2,2))
par(mar = c(4, 4, 2, 1))
plotdistcens(d, NPMLE = FALSE, main = "")
cdfcompcens(f)
qqcompcens(f)
ppcompcens(f)
```

## How to use of these new goodness-of-fit plots ?

Example of code:
```{r, eval = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
# Plot of the NPMLE CDF on censored data
plotdistcens(d)
# Two MLE fits
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
# Three goodness-of-fit plots for one fit
plot(fitsfn)
# Goodness-of-fit plots for one or more fits
cdfcompcens(list(fitsfn,fitsfl))
qqcompcens(list(fitsfn,fitsfl))
ppcompcens(list(fitsfn,fitsfl))
```

## Other recent improvements of fitdistrplus

**Version 1.0-8**

* add of an optional use of ggplot2 in cdfcomp(), denscomp(), qqcomp() and ppcomp().

**Version 1.0-10**

* Improvement of goodness-of-fit plots for discrete distributions in denscomp().
* Add of new default starting values for distributions in actuar.

**Version 1.0-11**

* add of an optional use of ggplot2 in cdfcompcens(), denscompcens() and ppcompcens().

## References
* Turnbull BW (1974). Nonparametric estimation of a survivorship function with doubly censored data. Journal of American Statistical Association, 69, 169-173.
* Gentleman, R., & Geyer, C. J. (1994). Maximum likelihood for interval censored data: Consistency and computation. Biometrika, 81(3), 618-623.
* Wang, Y. (2008). Dimension-reduced nonparametric maximum likelihood computation for interval-censored data. Computational Statistics & Data Analysis, 52(5), 2388-2402.
* Wang, Y., & Taylor, S. M. (2013). Efficient computation of nonparametric survival functions via a hierarchical mixture formulation. Statistics and Computing, 23(6), 713-725.
* Wang, Y., & Fani, S. (2018). Nonparametric maximum likelihood computation of a U-shaped hazard function. Statistics and Computing, 28(1), 187-200.

## 
\begin{center}
\begin{Large}
Thank you for your attention !

We are waiting for your feedback on these new tools.
\end{Large}
\end{center}